---
title: "GSS challange More Data"
author: "Chi Nguyen"
date: "summer2024"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

This is a type of categorical problem, with at least one problem with multiple level. To account for this we will use a Fisher test.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

## Now with The Most data (fewer variables gives more data points)

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In this graph, the first slide display the number of individuals categorized by races, white, black anf other. The white race had the most with over 1500 reports and the black and other race had less than 250 reports
In the last display the chart shows the percentage of the races on favors capital punishment. Each category was compared to each other, the white was compared to support versus the opposition of capital punishment and was in favor of capital punishment by approximately 64%. The black race was in opposition, with a much lower percentage of round 57%. The other race was in favor of capital punishment with a near report for its opposition, the support percentage was around 55%. 


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->




## Now with More data

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```

The sumary display that 65% of the white race was reported, 42% of the black race, and 55% of the other race were all in favor of capital punishment. This is compared to 34% of the white race in opposition of capital punishment , 57% of the black race and 44% of the other race 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```

The sumary displays that 65% of the white race was reported, 42% of the black race, and 55% of the other race were all in favor of capital punishment. This is compared to 34% of the white race in opposition of capital punishment , 57% of the black race and 44% of the other race 


## More data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
```

```{r}
fisherXY(mydata$race, mydata$cappun, printAll = "TRUE")
```

 In the inferential data, we ran a CHI square test and a Fisher test to get multiple results of each comparison. The first test, the CHI squared shared that the probability it pulled was equal to less than .000000000000000022. This number is too small.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

To conclude this report we aare looking at the complied data and asking ourselves if it matches the hypothesis prediction. In this area of data we found that the number of people belonging to the white race category were overwhelming the data for the other two races. 
